Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-07-02 edubots_surveys.Rnw
integrates Olia' s contribution refs #24

%%% Load packages
\usepackage{amsthm,amsmath}

@@

A chi-square test of independence was performed to examine the relation between instructors’ discipline, sector, and gender and any potential changes that occurred to the use of messaging apps due to the covid-19 pandemic. The relation between the latter variable and the instructors’ sector was significant, X2 (4, N = 282) = 9.598, p =0.048. The frequencies cross tabulated are given in Table 2. In effect, this finding indicates that how teachers responded to the use of messaging apps during the pandemic and in particular, whether they changed their habits in the use of apps for teaching purposes, was related to their sector (university vs vocational). The relationships between changes in the use of messaging apps, due to the covid-19 pandemic, and the instructors’ discipline, X2 (24, N = 282) = 44.856, p =0.006, as well as the gender, X2 (8, N = 282) = 16.249, p= 0.039 were also significant.  
% Olia - Fran, please also add here tables 1 and 2, as in the attachment that I have sent you. And check for repetitions in the following paragraph that you added. thank you.
The descriptive analysis in this case is done using a Chi-square test. For the different variables considered, some statistical significance was found for Gender, Sector, and academic Discipline. The values found with the test are: $p-value=0.0034$ for Gender, $p-value=0.0199$ for Sector, and $p-value=0.0211$ for Discipline.
The first column in Fig. \ref{figure:postcovid_chisq} shows the Pearson's residuals of the test for these three variables. Residuals show the difference between the observed and the expected values, with large residuals showing that variables are not truly independent. In this case, blue shows positive contribution and red negative contribution to the test, and saturation shows how large is this contribution in contrast to the expected value by chance. Tables on the right side show adjusted standardized residuals that according to \cite{agresti2013categorical}, if greater than +/-2 for cases with few cells, indicate lack of fit of $H_0$ (bolded in our table). Also, the table shows the percentage of contribution to the test of each cell.

@@

For answering RQ2, teachers were provided with a list of different potential chatbot functionalities (use cases) and were requested to respond on whether each given use case would be useful in their classes.  The findings are summarized in the table that follows. 
% Fran pleas add table 3 here, as in the attachment I sent you, along with the given caption. 
For the findings it occurs that the most favourable use case for chatbots in Higher Education and vocational training is their use for the facilitation of an agenda formation (171 positive response, 60.6%), followed by the FAQs use case (148 positive responses, 52.5%) and the sharing class material use case (136; 48.2%). A chi-square test of independence was performed to examine the relation among participants’ preferences for particular chatbot use cases. The relation between agenda and the FAQs use case was significant, X2 (1, N = 282) = 10.467, p =0.001. The frequencies cross tabulated are given in Table 4. Out of the 171 teachers who consider useful the use of chatbots for agenda preparation in the class, 103 also consider useful chatbots’ use for FAQs.  
% Fran pleas add table 4 here, as in the attachment I sent you, along with the given caption. 
Answer to these questions are plotted in Figure \ref{figure:chatbots_types} grouped by gender, sector and experience. As it was a multiple choice question, the counts are over the total number of people that answered. One of the most interesting questions that can arise from the plots, and the should require further study, is: Are our professors really aware of the possibilities of chatbots in the classroom? The reason to ask ourselves this question is due to the fact of the small number of responses for other uses in comparison with the other approaches proposed. Another way to loook about it is that, unless you provide the specific instructions on how to use a tool, the teaching staff is already too busy or burn out to think about innovation.

@@

For answering RQ3, teachers were provided with a list of different kind of interactions that may take place among students and between students and the teacher, with the use of messaging apps. The findings are summarized in the table that follows. 
% Fran please add table 5, as per attachment, along with the given caption.

@@

\subsection{RQ4 - What kind of interaction media features do teachers value the most?}

@@

A total of 205 teachers responded to the second survey, 124 males (60.5%), 65 females (31.7%) and 16 individuals who chose not to identify their gender (7.8%). The majority of the teachers were aged 45-55 years old (70; 34.1%), followed by individuals aged 35-45 (67, 32.7%), and 25-35 (42; 20.5%), while the minority was aged >55 years old (26; 12.7%). In terms of teaching experience, 59 teachers (28.8%) had an experience of 16-25 years, followed by teachers who had a teaching experience of 06-15 years (51; 24.9%), 0-5 years (51; 24.9%) and more than 25 years of experience (44; 21.5%). The teachers were asked about what kind of interaction media features they value the most; the frequencies and percentages of their responses are given in the table that follows. 
% Fran, please add table 6 here, as per attachment, along with the given caption.

A chi-square test of independence was performed to examine the potential relationship among the various features. The statistically significant correlations are given in the table that follows.  
% Fran, please add table 7 here, as per attachment, along with the given caption.

Finally, A two-step cluster analysis was conducted, with a log-likelihood distance measure adopted, to explore how the cases could be grouped based on their preferences of specific interaction media features, and which features have larger predictor importance for the clustering. Two clusters occurred from the analysis: cluster 1 (n=78, 60.9%) was comprise of instructors who did not value interaction media pluggability, connectivity, official formation, but valued interaction media analytics, familiarity, officiality, while cluster 2 (n=50, 39.1%) was comprise of instructors from which the majority did not value media analytics, official formation, familiarity, but valued pluggability, connectivity, and officiality. Interaction media pluggability appears to have the most important predictor importance in the clustering of cases (predictor importance = 1.0), whilst media unidirectionality was the least important factor (predictor importance = 0.04). The cluster quality is fair, but not a good one (silhouette measure of cohesion and separation). 
% Fran, please add figure 1 here as per attachment, along with the given caption.

@@

 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly.

\end{itemize}
% TODO A chi-square test of independence was performed to examine the potential relationship among the various features. The statistically significant correlations are given in the table that follows.

>>>>

%%% Load packages
\usepackage{amsthm,amsmath}
\usepackage{multirow}

@@

A chi-square test of independence was performed to examine the relation between instructors’ discipline, sector, and gender and any potential changes that occurred to the use of messaging apps due to the covid-19 pandemic. The relation between the latter variable and the instructors’ sector was significant, $X^{2}_{(4, N = 282)} = 9.598$, $p =0.048$. The frequencies cross tabulated are given in Table \ref{tab:sector*covid}. In effect, this finding indicates that how teachers responded to the use of messaging apps during the pandemic and in particular, whether they changed their habits in the use of apps for teaching purposes, was related to their sector (university vs vocational). The relationships between changes in the use of messaging apps, due to the covid-19 pandemic, and the instructors’ discipline, $X^{2}_{(24, N = 282)} = 44.856$, $p =0.006$, as well as the gender, $X^{2}_{(8, N = 282)} = 16.249$, $p= 0.039$ were also significant.

%TODO: This table is not referenced in the text
\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Messaging App} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5} 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
Telegram & 62 & 22.0 & 220 & 78.0 \\
WhatsApp & 124 & 44.0 & 158 & 56.0 \\
Slack & 15 & 5.3 & 267 & 94.7 \\
Other & 60 & 21.3 & 222 & 78.7 \\
\begin{tabular}[c]{@{}l@{}}Provided by the Academic\\ Institution\end{tabular} & 159 & 56.4 & 123 & 43.6 \\
None & 19 & 6.7 & 263 & 93.3 \\ \hline
\end{tabular}
\caption{Use of messaging apps in class}
\label{tab:use_messapps}
\end{table}

\begin{table}[h!tbp]
\begin{tabular}{llccc}
\hline
 & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Sector} & Total \\ \cline{3-5} 
 & \multicolumn{1}{c}{} & \begin{tabular}[c]{@{}c@{}}University \\ (f)\end{tabular} & \begin{tabular}[c]{@{}c@{}}Vocational\\ (f)\end{tabular} &  \\ \cline{2-5} 
\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Post-covid\\ changes\end{tabular}} & \begin{tabular}[c]{@{}l@{}}Yes, I use a safer messaging \\ app now\end{tabular} & 27 & 13 & 40 \\
 & \begin{tabular}[c]{@{}l@{}}Yes, I didn’t use any \\ messaging app before, but \\ do use it now\end{tabular} & 31 & 13 & 44 \\
 & \begin{tabular}[c]{@{}l@{}}No, I did use it before and \\ use it now\end{tabular} & 105 & 60 & 165 \\
 & \begin{tabular}[c]{@{}l@{}}No, I didn’t use it before and \\ don’t use it now\end{tabular} & 22 & 2 & 24 \\
Total &  & 193 & 89 & 282 \\ \hline
\end{tabular}
\caption{Sector * post-covid changes crosstabulation}
\label{tab:sector*covid}
\end{table}

% Olia - TODO: check for repetitions in the following paragraph that you added. thank you.
The descriptive analysis in this case is done using a Chi-square test. For the different variables considered, some statistical significance was found for Gender, Sector, and academic Discipline. The values found with the test are: $p-value=0.0034$ for Gender, $p-value=0.0199$ for Sector, and $p-value=0.0211$ for Discipline.
The first column in Fig. \ref{figure:postcovid_chisq} shows the Pearson's residuals of the test for these three variables. Residuals show the difference between the observed and the expected values, with large residuals showing that variables are not truly independent. In this case, blue shows positive contribution and red negative contribution to the test, and saturation shows how large is this contribution in contrast to the expected value by chance. Tables on the right side show adjusted standardized residuals that according to \cite{agresti2013categorical}, if greater than +/-2 for cases with few cells, indicate lack of fit of $H_0$ (bolded in our table). Also, the table shows the percentage of contribution to the test of each cell.

@@

For answering RQ2, teachers were provided with a list of different potential chatbot functionalities (use cases) and were requested to respond on whether each given use case would be useful in their classes. The findings are summarized in Table \ref{tab:chatbottypes}.

\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Chatbot use cases} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5} 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
Answering to students' FAQs & 148 & 52.5 & 134 & 47.5 \\
Assigning student grades & 113 & 40.1 & 169 & 59.9 \\
Facilitating agenda information & 171 & 60.6 & 111 & 39.4 \\
Sharing class materials & 136 & 48.2 & 146 & 51.8 \\
Others & 25 & 8.9 & 257 & 91.1 \\ \hline
\end{tabular}
\caption{Perceived useful chatbot use cases}
\label{tab:chatbottypes}
\end{table}

For the findings it occurs that the most favourable use case for chatbots in Higher Education and vocational training is their use for the facilitation of an agenda formation (171 positive response, 60.6\%), followed by the FAQs use case (148 positive responses, 52.5\%) and the sharing class material use case (136; 48.2\%). A chi-square test of independence was performed to examine the relation among participants’ preferences for particular chatbot use cases. The relation between agenda and the FAQs use case was significant, $X^{2}_{(1, N = 282)} = 10.467$, $p =0.001$. The frequencies cross tabulated are given in Table \ref{tab:agenda*faq}. Out of the 171 teachers who consider useful the use of chatbots for agenda preparation in the class, 103 also consider useful chatbots’ use for FAQs.

\begin{table}[h!tbp]
\begin{tabular}{llccc}
\hline
 & \multicolumn{1}{c}{} & \multicolumn{2}{c}{FAQs} & Total \\ \cline{3-5} 
 & \multicolumn{1}{c}{} & Yes & No &  \\ \cline{2-5} 
\multirow{2}{*}{Agenda} & Yes & 103 & 68 & 171 \\
 & No & 45 & 66 & 111 \\
Total &  & 148 & 134 & 282 \\ \hline
\end{tabular}
\caption{Agenda use case * FAQs use case crosstabulation }
\label{tab:agenda*faq}
\end{table}

Answer to these questions are plotted in Figure \ref{figure:chatbots_types} grouped by gender, sector and experience. As it was a multiple choice question, the counts are over the total number of people that answered. One of the most interesting questions that can arise from the plots, and the should require further study, is: Are our professors really aware of the possibilities of chatbots in the classroom? The reason to ask ourselves this question is due to the fact of the small number of responses for other uses in comparison with the other approaches proposed. Another way to loook about it is that, unless you provide the specific instructions on how to use a tool, the teaching staff is already too busy or burn out to think about innovation.

@@

For answering RQ3, teachers were provided with a list of different kind of interactions that may take place among students and between students and the teacher, with the use of messaging apps. The findings are summarized in Table \ref{tab:kindinteractions}.

\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Kind of interactions} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5} 
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
\begin{tabular}[c]{@{}l@{}}Chat interactions among \\ students in the same course\end{tabular} & 171 & 60.6 & 110 & 39.0 \\
\begin{tabular}[c]{@{}l@{}}Chat interactions among students, \\ teachers, and the faculty\end{tabular} & 13 & 4.6 & 269 & 95.4 \\
\begin{tabular}[c]{@{}l@{}}Chat interactions among students \\ of the same study year and teachers\end{tabular} & 51 & 18.1 & 231 & 81.9 \\
\begin{tabular}[c]{@{}l@{}}Teacher not being part of the \\ interaction\end{tabular} & 113 & 40.1 & 169 & 59.9 \\ \hline
\end{tabular}
\caption{Kind of interactions preferred}
\label{tab:kindinteractions}
\end{table}

@@

%\subsection{RQ4 - What kind of interaction media features do teachers value the most?}

@@

%A total of 205 teachers responded to the second survey, 124 males (60.5\%), 65 females (31.7\%) and 16 individuals who chose not to identify their gender (7.8\%). The majority of the teachers were aged 45-55 years old (70; 34.1\%), followed by individuals aged 35-45 (67, 32.7\%), and 25-35 (42; 20.5\%), while the minority was aged >55 years old (26; 12.7\%). In terms of teaching experience, 59 teachers (28.8\%) had an experience of 16-25 years, followed by teachers who had a teaching experience of 06-15 years (51; 24.9\%), 0-5 years (51; 24.9\%) and more than 25 years of experience (44; 21.5\%). The teachers were asked about what kind of interaction media features they value the most; the frequencies and percentages of their responses are given in the table that follows.

@@

 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly. 
\end{itemize}

A chi-square test of independence was performed to examine the potential relationship among the various features. The statistically significant correlations are given in Table \ref{tab:interactionfeatures}.

\begin{table}[h!tbp]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{llllllllllll}
\hline
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Analytics} & Connectivity & Familiarity & Hidden phone & Horizontallity & \begin{tabular}[c]{@{}l@{}}Official\\  formation\end{tabular} & Pluggability & Sustainability & Unidirectionality & Officiality & Synchrony \\ \hline
Analytics & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{$X^{2}_{(1, N = 205)} = 12.172^{***}$} & $X^{2}_{(1, N = 205)} = 4.855^{*}$ & ns & ns & $X^{2}_{(1, N = 205)} = 4.144^{*}$ & \multicolumn{1}{c}{$X^{2}_{(1, N = 205)} = 12.490^{***}$} & $X^{2}_{(1, N = 205)} = 4.314^{*}$ & ns & ns & ns \\
Connectivity & \multicolumn{1}{c}{} & \multicolumn{1}{c}{-} & ns & ns & $X^{2}_{(1, N = 205)} = 5.971^{***}$ & ns & \multicolumn{1}{c}{$X^{2}_{(1, N = 205)} = 16.286^{***}$} & ns & ns & ns & ns \\
Familiarity &  &  & - & ns & ns & ns & ns & ns & ns & ns & ns \\
Hidden phone &  &  &  & - & ns & $X^{2}_{(1, N = 205)} = 5.678^{*}$ & $X^{2}_{(1, N = 205)} = 4.683^{*}$ & ns & $X^{2}_{(1, N = 205)} = 4.515^{*}$ & ns & ns \\
Horizontallity &  &  &  &  & - & ns & ns & ns & ns & $X^{2}_{(1, N = 205)} = 5.453^{*}$ & ns \\
\begin{tabular}[c]{@{}l@{}}Official\\  formation\end{tabular} &  &  &  &  &  & - & $X^{2}_{(1, N = 205)} = 8.163^{**}$ & ns & ns & ns & ns \\
Pluggability &  &  &  &  &  &  & - & ns & ns & ns & ns \\
Sustainability &  &  &  &  &  &  &  & - & ns & ns & ns \\
Unidirectionality &  &  &  &  &  &  &  &  & - & ns & ns \\
Officiality &  &  &  &  &  &  &  &  &  & - & ns \\
Synchrony & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} &  &  &  &  & \multicolumn{1}{c}{} &  &  &  & - \\ \hline
\end{tabular}}
\caption{Interaction media features analysis ($^{***}p<0.001$, $^{**}p<0.01$, $^{*}p<0.05$).}
\label{tab:interactionfeatures}
\end{table}

Finally, A two-step cluster analysis was conducted, with a log-likelihood distance measure adopted, to explore how the cases could be grouped based on their preferences of specific interaction media features, and which features have larger predictor importance for the clustering. Two clusters occurred from the analysis: cluster 1 (n=78, 60.9\%) was comprise of instructors who did not value interaction media pluggability, connectivity, official formation, but valued interaction media analytics, familiarity, officiality, while cluster 2 (n=50, 39.1\%) was comprise of instructors from which the majority did not value media analytics, official formation, familiarity, but valued pluggability, connectivity, and officiality. Interaction media pluggability appears to have the most important predictor importance in the clustering of cases (predictor importance = 1.0), whilst media unidirectionality was the least important factor (predictor importance = 0.04). The cluster quality is fair, but not a good one (silhouette measure of cohesion and separation).

% Fran, please add figure 1 here as per attachment, along with the given caption.
